Deriving Multi-Headed Projective Dependency Parses \\ from Link Grammar Parses

### 2014-06-03 doc/LinksToDAG.tex
minor edits

\begin{abstract}
Under multi-headed dependency grammar, a parse is a directed acyclic graph rather than a tree.  Such formalisms can be more syntactically and semantically expressive.  However, 
it is hard to train, test, or improve multi-headed parsers because few multi-headed corpora exist, particularly for the projective case.
% OLD
% investigate the benefit of such parsers or to work on making them faster or more accurate.  
To help fill this gap, we observe that link grammar already produces {\em undirected} projective graphs.  
% OLD
% ... produces parses that are similar to multi-headed projective dependency parses, except that the links are undirected. 
We use Integer Linear Programming to assign consistent directions to the labeled links in a corpus of several thousand parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English as well as Russian and other languages.  We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses. The resulting parses in English appear reasonably linguistically plausible, although their style does not match CoNLL-style parses of the same sentences; we discuss the differences.  
\end{abstract}

@@

Our goal in this paper is to recover these implicit directions by global optimization.  We seek a fixed mapping from labels to directions such that we can interpret link grammar parses as directed dependency parses that satisfy all of our constraints.

@@

\section{Experiments and Results}
\begin{figure*}[ht!]
  \centering
  \input{figure/sentences.tikz}
  \caption{\small Example Parses: The top blue arcs show CoNLL dependency parses; the bottom red arcs show oriented link grammar parses. See Appendix~\ref{app:parses} for formatting details and more examples.}
  \label{fig:parses}
\end{figure*}

@@

As noted earlier, the resulting English corpus uses an multi-headed annotation scheme that is somewhat different from the CoNLL annotations.  Differences are tabulated in Appendix~\ref{app:linktypes}, and the actual parses are contrasted in Figure~\ref{fig:parses} and Appendix~\ref{app:parses}.

@@

\NoteJE{improve table formatting}

@@

Below we display the first 100 English sentences whose widths do not exceed the width of this page.  The blue top half shows the CoNLL tagging and dependency parse.  The red bottom half shows the link grammar's tagging and our orientation of the link parse.  
% The dashes in the link grammar part of speech tag annotations mean that the link parser did not return one.
Words that were unknown to the link parser are flagged with \texttt{[!]}.

>>>>

\begin{abstract}
Under multi-headed dependency grammar, a parse is a directed acyclic graph rather than a tree.  Such formalisms can be more syntactically and semantically expressive.  However, 
it is hard to train, test, or improve multi-headed parsers because few multi-headed corpora exist, particularly for the projective case.
% OLD
% investigate the benefit of such parsers or to work on making them faster or more accurate.  
To help fill this gap, we observe that link grammar already produces {\em undirected} projective graphs.  
% OLD
% ... produces parses that are similar to multi-headed projective dependency parses, except that the links are undirected. 
We use Integer Linear Programming to assign consistent directions to the labeled links in a corpus of several thousand parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English as well as Russian and other languages.  We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses. The resulting parses in English appear reasonably linguistically plausible, though differing in style from CoNLL-style parses of the same sentences; we discuss the differences.  
\end{abstract}

@@

Our goal in this paper is to recover these implicit directions by global optimization.  We seek a fixed mapping from labels to directions such that link grammar parses become directed dependency parses that satisfy all of our constraints.

@@

\section{Experiments and Results}
\begin{figure*}[ht!]
  \centering
  \input{figure/sentences.tikz}
  \caption{\small Example Parses: The blue upper edges show CoNLL dependency parses; the red lower edges show oriented link grammar parses. Vertical edges are those with parent 0.  See Appendix~\ref{app:parses} for formatting details and more examples.}
  \label{fig:parses}
\end{figure*}

@@

The resulting English corpus uses an multi-headed annotation scheme that is somewhat different from the CoNLL annotations.  Differences are tabulated in Appendix~\ref{app:linktypes}, and the actual parses are contrasted in Figure~\ref{fig:parses} and Appendix~\ref{app:parses}.

@@

\NoteJE{improve table formatting.  Numeric justification; longtable; remove most hlines.}

@@

Below we display the first 100 English sentences whose widths do not exceed the width of this page.  The blue upper half shows the CoNLL tagging and dependency parse.  The red lower half shows the link grammar's tagging and our orientation of the link parse.  
% The dashes in the link grammar part of speech tag annotations mean that the link parser did not return one.
A vertical edge to word token $v$ denotes an edge $0 \rightarrow v$.
Words that were unknown to the link parser are flagged with \texttt{[!]}.
